A finite-state morphological transducer for Kyrgyz

### 2012-03-13 paper/paper.tex
more tweaking
update stems table

\begin{table}[htbp]
	\caption{Examples comparing desonorisation of \{N\} and \{L\}}\label{exdesonorisation}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/алма-NI/ & [алманы] & \eng{apple--\gmk{ACC}} \\
			/сыр-NI/ & [сырды] & \eng{secret--\gmk{ACC}} \\
			/каз-NI/ & [казды] & \eng{goose--\gmk{ACC}} \\
			/баш-NI/ & [башты] & \eng{head--\gmk{ACC}} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{2em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/алма-LAр/ & [алмалар] & \eng{apple--\gmk{PL}} \\
			/сыр-LAр/ & [сырлар] & \eng{secret--\gmk{PL}} \\
			/каз-LAр/ & [каздар] & \eng{goose--\gmk{PL}} \\
			/баш-LAр/ & [баштар] & \eng{head--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
\end{table}

@@

Kyrgyz has a number of nouns that underlyingly end with a consonant cluster consisting of two liquids, but which are split by an epenthetic vowel syllable-finally.
\begin{table}[htbp]
	\caption{Examples of /рн/-final nouns surfacing with epenthetic vowel}\label{exrnvowel}
	\centering
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/мурн/ & [мурун] & \eng{nose} \\
			/мурн-LAр/ & [мурундар] & \eng{nose--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{2em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/орн/ & [орун] & \eng{place} \\
			/орн-LAр/ & [орундар] & \eng{places--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
\end{table}
However, when a vowel-initial suffix follows the noun, the epenthetic vowel is absent, and all rules of consonant unfaithfulness involving consonant clusters are enforced[?].
%FIXME: "[?]"
%FIXME: make better references to the tables
\begin{table}[htbp]
	\caption{Examples of /рн/-final nouns surfacing with no epenthetic vowel}\label{exrnnovowel}
	%\centering
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			%FIXME: should the glosses here and below be morpheme-by-morpheme for consistency, or is a more colloquial gloss okay?
			/мурн-(S)I/ & [мурду] & \eng{his/her/its nose} \\
			/мурн-Iм-DAн/ & [мурдуман] & \eng{from my nose} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{0.5em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/орн-(S)I/ & [орду] & \eng{his/her/its place} \\
			/орн-Iм-DAн/ & [ордуман] & \eng{from my place} \\
			\bottomrule
		\end{tabular}
	}
\end{table}

@@

\begin{table*}
	\centering
	\begin{tabular}{lr}
		\toprule
		Part of speech & Number of stems\\
		\midrule
		Noun & 4,731\\
		Adjective & 924\\
		Verb & 896 \\
		Proper noun & 745\\
		Adverb & 213\\
		Numeral & 50\\
		Postposition & 42\\
		Conjunction & 36\\
		Pronoun & 29\\
		Determiner & 27\\
		\midrule
		Total:     & 7,333\\
		\bottomrule
	\end{tabular}
	\caption{FIXME}\label{table:stems}
	%FIXME: are we putting captions above or below tables?  I like them better below, but I got the impression that there's some standard to have them above.  Maybe it was just UW being table Nazis for my MA thesis.
\end{table*}

>>>>

\begin{table}[htbp]
	\caption{Examples comparing desonorisation of \{N\} and \{L\}}\label{exdesonorisation}
\begin{center}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/алма-NI/ & [алманы] & \eng{apple--\gmk{ACC}} \\
			/сыр-NI/ & [сырды] & \eng{secret--\gmk{ACC}} \\
			/каз-NI/ & [казды] & \eng{goose--\gmk{ACC}} \\
			/баш-NI/ & [башты] & \eng{head--\gmk{ACC}} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{2em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/алма-LAр/ & [алмалар] & \eng{apple--\gmk{PL}} \\
			/сыр-LAр/ & [сырлар] & \eng{secret--\gmk{PL}} \\
			/каз-LAр/ & [каздар] & \eng{goose--\gmk{PL}} \\
			/баш-LAр/ & [баштар] & \eng{head--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
\end{center}
\end{table}

@@

Kyrgyz has a number of nouns that underlyingly end with a consonant cluster consisting of two liquids, but which are split by an epenthetic vowel syllable-finally.
\begin{table}[htbp]
	\caption{Examples of /рн/-final nouns surfacing with epenthetic vowel}\label{exrnvowel}
\begin{center}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/мурн/ & [мурун] & \eng{nose} \\
			/мурн-LAр/ & [мурундар] & \eng{nose--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{2em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/орн/ & [орун] & \eng{place} \\
			/орн-LAр/ & [орундар] & \eng{places--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
\end{center}
\end{table}
However, when a vowel-initial suffix follows the noun, the epenthetic vowel is absent, and all rules of consonant unfaithfulness involving consonant clusters are enforced[?].
%FIXME: "[?]"
%FIXME: make better references to the tables
\begin{table}[htbp]
	\caption{Examples of /рн/-final nouns surfacing with no epenthetic vowel}\label{exrnnovowel}
\begin{center}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			%FIXME: should the glosses here and below be morpheme-by-morpheme for consistency, or is a more colloquial gloss okay?
			/мурн-(S)I/ & [мурду] & \eng{his/her/its nose} \\
			/мурн-Iм-DAн/ & [мурдуман] & \eng{from my nose} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{0.5em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/орн-(S)I/ & [орду] & \eng{his/her/its place} \\
			/орн-Iм-DAн/ & [ордуман] & \eng{from my place} \\
			\bottomrule
		\end{tabular}
	}
\end{center}
\end{table}

@@

\begin{table*}
	\centering
	\begin{tabular}{lr}
		\toprule
		Part of speech & Number of stems\\
		\midrule
		Noun & 4,972\\
		Verb & 1,231 \\
		Adjective & 944\\
		Proper noun & 796\\
		Adverb & 295 \\
		Numeral & 63 \\
		Conjunction & 58 \\
		Postposition & 51 \\
		Pronoun & 29\\
		Determiner & 27\\
		\midrule
		Total: 8,466 & \\
		\bottomrule
	\end{tabular}
	\caption{Breakdown of number of stems in the lexicon by major part of speech. Minor categories, such as partice, modal and copula are left out.}\label{table:stems}
	%FIXME: are we putting captions above or below tables?  I like them better below, but I got the impression that there's some standard to have them above.  Maybe it was just UW being table Nazis for my MA thesis.
\end{table*}
